### 2016-06-24 Projects/DelCo_2014_2015/Reports/Hose_Stream_Tests/TeX_Files/DelCo_Hose_Stream_Tests.tex
Added abstract to hose stream report

% \chapter{List of Symbols}
% \begin{tabbing}
% \hspace{1.5in} \= \\
% ft \> foot \\
% GPM \> gallons per minute \\
% in \> inch \\
% m \> meter \\
% \end{tabbing}

@@

% ================
% = Introduction =
% ================
\chapter{Introduction}
\label{chap:intro}
\section{Background}
\label{sec:background}
One of the objectives of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to improve the safety and effectiveness of firefighters through improved knowledge of fire behavior and examination of how different firefighting tactics affect the fire environment. NIST has conducted a significant amount of research examining how ventilation affects the growth and spread of fire within structures and how the air flow to the fire may be controlled to limit or delay the growth of the fire~\cite{madrzykowski2009fire,kerber2009fire}. The studies have resulted in guidance to the fire service regarding ventilation tactics. However, ventilation tactics alone will not extinguish the fire; fire suppression with hose streams is also needed.

@@

Four types of hose streams and four nozzle movement patterns patterns were studied as part of the experiments described in this report. Three of the hose stream patterns studied were from a combination nozzle: straight stream, narrow fog stream, and wide fog stream. Water was flowed into the structure by having the hose in a static, or fixed, position; by moving the hoseline left-to-right across the room in a sweeping motion; and by rotating the hoseline in both the clockwise and counterclockwise directions. Additionally, experiments comparing the differences in impact on air flows in a structure from a combination nozzle straight stream and a solid stream from a smooth bore nozzle equipped with a 1~in tip were conducted. Images of the four types of hose streams are presented in Fig.~\ref{fig:hose_streams}.

@@

\begin{table}[!ht]
\caption{Average gas velocity [$\pm18$\%] through interior doorway when flow path was fully established for Tests~6 and 7}
\begin{tabular}{lccccc}
\toprule
 & & \multicolumn{4}{c}{\underline{Test 6}}
\\
\textbf{Stream} & \textbf{Units} & 
\textbf{Fixed} & \textbf{Sweeping} & \textbf{Clockwise} & \begin{tabular}{@{}c@{}} \textbf{Counter} \\ \textbf{Clockwise} \\ \end{tabular}
\\ \midrule
\multirow{2}{*}{\textit{Straight}} & 
\small{m/s}   &  -0.2 $\pm0.03$   &   0.5 $\pm0.1$  &   0.7 $\pm0.1$   &  0.7 $\pm0.1$  \\ & 
\small{(mph)} &  (-0.4 $\pm0.1$)  &  (1.2 $\pm0.2$) &  (1.6 $\pm0.3$)  & (1.7 $\pm0.3$)
\\~\\
\multirow{2}{*}{\textit{Narrow Fog}} & 
\small{m/s}   &   0.4 $\pm0.1$   &   1.1 $\pm0.2$  &   0.9 $\pm0.2$   &  1.1 $\pm0.2$  \\ & 
\small{(mph)} &  (0.9 $\pm0.2$)  &  (2.4 $\pm0.4$) &  (1.9 $\pm0.3$)  & (2.4 $\pm0.4$)
\\~\\
\multirow{2}{*}{\textit{Wide Fog}} & 
\small{m/s}   &   1.0 $\pm0.2$   &   1.0 $\pm0.2$  &   1.0 $\pm0.2$   &  0.9 $\pm0.2$  \\ & 
\small{(mph)} &  (2.2 $\pm0.4$)  &  (2.1 $\pm0.4$) &  (2.2 $\pm0.4$)  & (2.0 $\pm0.4$) \\ 
\midrule
 & & \multicolumn{4}{c}{\underline{Test 7}}
\\
\textbf{Stream} & \textbf{Units} & 
\textbf{Fixed} & \textbf{Sweeping} & \textbf{Clockwise} & \begin{tabular}{@{}c@{}} \textbf{Counter} \\ \textbf{Clockwise} \\ \end{tabular}
\\ \midrule
\multirow{2}{*}{\textit{Straight}} & 
\small{m/s}   &  -0.2 $\pm0.04$   &   0.5 $\pm0.1$  &   0.7 $\pm0.1$   &  0.7 $\pm0.1$  \\ & 
\small{(mph)} &  (-0.7 $\pm0.1$)  &  (2.8 $\pm0.5$) &  (3.1 $\pm0.6$)  & (2.7 $\pm0.5$)
\\~\\
\multirow{2}{*}{\textit{Narrow Fog}} & 
\small{m/s}   &  -0.2 $\pm0.04$   &   0.5 $\pm0.1$  &   0.7 $\pm0.1$   &  0.7 $\pm0.1$  \\ & 
\small{(mph)} &  (-0.7 $\pm0.1$)  &  (2.8 $\pm0.5$) &  (3.1 $\pm0.6$)  & (2.7 $\pm0.5$)
\\~\\
\multirow{2}{*}{\textit{Wide Fog}} & 
\small{m/s}   &  -0.2 $\pm0.04$   &   0.5 $\pm0.1$  &   0.7 $\pm0.1$   &  0.7 $\pm0.1$  \\ & 
\small{(mph)} &  (-0.7 $\pm0.1$)  &  (2.8 $\pm0.5$) &  (3.1 $\pm0.6$)  & (2.7 $\pm0.5$) \\ 
\bottomrule
\end{tabular}
\label{table:west_hand_A10_avgs}
\end{table}
\FloatBarrier

>>>>

% \chapter{List of Symbols}
% \begin{tabbing}
% \hspace{1.5in} \= \\
% ft \> foot \\
% gpm \> gallons per minute \\
% in \> inch \\
% m \> meter \\
% \end{tabbing}

@@

\chapter*{\centering Abstract}
\pagenumbering{gobble}

Fire suppression tactics using hose streams can affect ventilation in a structure and may impact the movement of smoke and heat through a structure. Seven series of experiments were conducted to study the impact of different hose stream and nozzle movement pattern combinations on air movement within residential scale structures. The experiments studied four different hose streams: a straight stream, narrow fog stream, and a wide fog stream from a combination nozzle and a solid stream from a smooth bore nozzle. The streams were applied in a static, or fixed, position; by moving the hoseline left-to-right across the room in a sweeping motion; and by rotating the hoseline in both the clockwise and counterclockwise directions. Gas velocity was measured at different locations in the structure during the experiments. The wide fog stream caused the most air movement out of any of the tested streams, reaching a maximum of 2.6~m/s, followed by the narrow fog stream, and then the straight stream and solid stream from the smooth bore nozzle, which both caused approximately the same amount of air movement through the structure. It was discovered that the straight stream only caused significant air movement through the structure when it was applied in a moving pattern. Furthermore, there were no statistically significant differences in the average measured air velocity between the clockwise and counterclockwise nozzle movement patterns. It was determined that the type of hose stream and manner in which it is applied dictates the extent to which a stream impacts the ventilation of a structure. To better quantify the impact of such air flows on the fire environment, additional fire experiments need to be conducted with structure fires in a controlled environment.

% ================
% = Introduction =
% ================
\chapter{Introduction}
\label{chap:intro}
\section{Background}
\label{sec:background}

One of the objectives of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to improve the safety and effectiveness of firefighters through improved knowledge of fire behavior and examination of how different firefighting tactics affect the fire environment. NIST has conducted a significant amount of research examining how ventilation affects the growth and spread of fire within structures and how the air flow to the fire may be controlled to limit or delay the growth of the fire~\cite{madrzykowski2009fire,kerber2009fire}. The studies have resulted in guidance to the fire service regarding ventilation tactics. However, ventilation tactics alone will not extinguish the fire; fire suppression with hose streams is also needed.

@@

Four types of hose streams and four nozzle movement patterns were studied as part of the experiments described in this report. Three of the hose stream patterns studied were from a combination nozzle: straight stream, narrow fog stream, and wide fog stream. Water was flowed into the structure by having the hose in a static, or fixed, position; by moving the hoseline left-to-right across the room in a sweeping motion; and by rotating the hoseline in both the clockwise and counterclockwise directions. Additionally, experiments comparing the differences in impact on air flows in a structure from a combination nozzle straight stream and a solid stream from a smooth bore nozzle equipped with a 1~in tip were conducted. Images of the four types of hose streams are presented in Fig.~\ref{fig:hose_streams}.

@@

\begin{table}[!ht]
\caption{Average gas velocity [$\pm18$\%] through interior doorway when flow path was fully established for Tests~6 and 7}
\begin{tabular}{lccccc}
\toprule
 & & \multicolumn{4}{c}{\underline{Test 6}}
\\
\textbf{Stream} & \textbf{Units} & 
\textbf{Fixed} & \textbf{Sweeping} & \textbf{Clockwise} & \begin{tabular}{@{}c@{}} \textbf{Counter} \\ \textbf{Clockwise} \\ \end{tabular}
\\ \midrule
\multirow{2}{*}{\textit{Straight}} & 
\small{m/s}   &  -0.2 $\pm0.03$   &   0.5 $\pm0.1$  &   0.7 $\pm0.1$   &  0.7 $\pm0.1$  \\ & 
\small{(mph)} &  (-0.4 $\pm0.1$)  &  (1.2 $\pm0.2$) &  (1.6 $\pm0.3$)  & (1.7 $\pm0.3$)
\\~\\
\multirow{2}{*}{\textit{Narrow Fog}} & 
\small{m/s}   &   0.4 $\pm0.1$   &   1.1 $\pm0.2$  &   0.9 $\pm0.2$   &  1.1 $\pm0.2$  \\ & 
\small{(mph)} &  (0.9 $\pm0.2$)  &  (2.4 $\pm0.4$) &  (1.9 $\pm0.3$)  & (2.4 $\pm0.4$)
\\~\\
\multirow{2}{*}{\textit{Wide Fog}} & 
\small{m/s}   &   1.0 $\pm0.2$   &   1.0 $\pm0.2$  &   1.0 $\pm0.2$   &  0.9 $\pm0.2$  \\ & 
\small{(mph)} &  (2.2 $\pm0.4$)  &  (2.1 $\pm0.4$) &  (2.2 $\pm0.4$)  & (2.0 $\pm0.4$) \\ 
\midrule
 & & \multicolumn{4}{c}{\underline{Test 7}}
\\
\textbf{Stream} & \textbf{Units} & 
\textbf{Fixed} & \textbf{Sweeping} & \textbf{Clockwise} & \begin{tabular}{@{}c@{}} \textbf{Counter} \\ \textbf{Clockwise} \\ \end{tabular}
\\ \midrule
\multirow{2}{*}{\textit{Straight}} & 
\small{m/s}   &   0.0 $\pm0.01$    &   0.5 $\pm0.1$  &   0.7 $\pm0.1$   &  0.7 $\pm0.1$  \\ & 
\small{(mph)} &  (-0.1 $\pm0.02$)  &  (1.2 $\pm0.2$) &  (1.6 $\pm0.3$)  & (1.5 $\pm0.3$)
\\~\\
\multirow{2}{*}{\textit{Narrow Fog}} & 
\small{m/s}   &   1.3 $\pm0.2$   &   1.2 $\pm0.2$  &   1.2 $\pm0.2$   &  1.3 $\pm0.2$  \\ & 
\small{(mph)} &  (2.9 $\pm0.5$)  &  (2.8 $\pm0.5$) &  (2.7 $\pm0.5$)  & (2.9 $\pm0.5$)
\\~\\
\multirow{2}{*}{\textit{Wide Fog}} & 
\small{m/s}   &   1.0 $\pm0.2$   &   0.7 $\pm0.1$  &   0.9 $\pm0.2$   &  1.0 $\pm0.2$  \\ & 
\small{(mph)} &  (2.2 $\pm0.4$)  &  (1.6 $\pm0.3$) &  (2.1 $\pm0.4$)  & (2.2 $\pm0.4$) \\ 
\bottomrule
\end{tabular}
\label{table:west_hand_A10_avgs}
\end{table}
\FloatBarrier
